---
title: "Covid91 vaccine study Final2020"
author: "Jordan Rose"
date: "Nov. 11th, 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

# guys
```{r}
guysData <- filter(FinalData,sex=="guy")
```
```{r}
barchartGC(~infected + treatment,data=guysData)
barchartGC(~infected + treatment,data=guysData, type="percent")

```

According to this graph we can say that the group of guys taking the drug were more likely to have caught COVID than the guys taking the placebo. This chart illustrates that the guys who were taking the drug were actually 20 times more likely to be infected than the guys taking the placebo. Therefor we can state that this was not a good drug in terms of preventing the infection of COVID.

### Numerical Summaries for Guys  

```{r}
table1 <- xtabs(~infected + treatment, data=guysData)
rowPerc(table1)
colPerc(table1)
```

It looks like 58% of the guys who got infected have taken the drug, and only 42% of the guys have taken the placebo. 

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Gals
```{r}
galData <- filter(FinalData,sex=="gal")
```
```{r}
barchartGC(~infected + treatment,data=galData)
barchartGC(~infected + treatment,data=galData, type="percent")

```

The drug appears to have been more effective in terms of avoiding COVID infection for the gals than it was for the guys. We can support this statement by looking at the results shown above. According to said data it appears that almost 35 more gals who were taking the placebo were being infected as compared to the 35 gals who were taking the drug. The data clearly illustrates that the gals who were taking the drug were less likely to catch  as the ones taking the placebo therefore the drug is more effective for the gals. 

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

# Drug Users
```{r}
DrugData <- filter(FinalData,DrugUser=="yes")
```
```{r}
barchartGC(~infected + treatment,data=DrugData)
barchartGC(~infected + treatment,data=DrugData, type="percent")

```

The drug appears to be working as well for the drug users as it did for the gals. The graph illustrtates and almost identical depiction for the drug users taking the drug as it showed for the gals taking the drug. More specifically we can say that 65 drug users who were taking the placebo were actually the ones who caught COVID while the drug users who were taking the drug were half as likely to catch it. 

# LGBTQ
```{r}
lgbtqData <- filter(FinalData,LGBTQ=="gay")
```
```{r}
barchartGC(~infected + treatment,data=lgbtqData)
barchartGC(~infected + treatment,data=lgbtqData, type="percent")

```

According to this graph more than twice the people whom have taken the drug have ended up catching COVID. Therefore this drug is not beneficial for the lgbtq community in terms of preventing the COVID infection. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
