---
title: "Covid91 vaccine study Final2020"
author: "Jordan Rose"
date: "Nov. 11th, 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

# guys
```{r}
guysData <- filter(FinalData,sex=="guy")
```
```{r}
barchartGC(~infected + treatment,data=guysData)
barchartGC(~infected + treatment,data=guysData, type="percent")

```

According to this graph we can say that the group of guys taking the drug were more likely to have caught COVID than the guys taking the placebo. This chart illustrates that the guys who were taking the drug were actually 20 times more likely to be infected than the guys taking the placebo. Therefor we can state that this was not a good drug in terms of preventing the infection of COVID.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Gals
```{r}
galData <- filter(FinalData,sex=="gal")
```

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

# Drug Users
```{r}
DrugData <- filter(FinalData,DrugUser=="yes")
```

# LGBTQ
```{r}
lgbtqData <- filter(FinalData,LGBTQ=="gay")
```

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
